### 2020-04-21 survival.tex
more cox

If $T$ is the random variable of a GPU fail time, then its cumulative
distribution function $F(t) = Pr\{T < t\}$ gives the probability that
a GPU fails by duration $t$. The survival function is its complement
\begin{displaymath}
  S(t) = Pr\{T \geq t\} = 1 - F(t).
\end{displaymath}
It is the probability of being operational at duration $t$.  We use
the R packages \pkg{survival} and \pkg{survminer} for the KM analysis,
which is reported in Figure~\ref{fig:km-all-cage-node}. Within each
{\em batch}, separate survival curves are computed for each {\em cage}
by {\em node} combination. Along with the survival curve estimate,
this analysis provides 95\% confidence region for survival probability
shaded around the curves.

@@

CPH is considered a semi-parametric model as there are no assumptions
about the shape of the baseline hazard function. But its main
assumption is that the hazards are proportional. There are a number of
tests for this, including graphical diagnostics in the \pkg{survminer}
package as well as checking that survival curves for categorical
covariates do not cross. We ran these diagnostics and concluded that
the hazards for our location categories are approximately
proportional. However, survival functions partitioned on {\em batch}
do cross and so we fit the model to the new and the old batches
separately. The results are presented in Figures~\ref{fig:cox-old}
and~\ref{fig:cox-new}.
\begin{figure}
  \centering
  \includegraphics[width=0.8\columnwidth]{figs/cox_o001.pdf}
  \caption{GPU hazard ratios from Cox regression model on {\tt old}
    batch.}
  \label{fig:cox-old}
\end{figure}
\begin{figure}
  \centering
  \includegraphics[width=0.8\columnwidth]{figs/cox_n001.pdf}
  \caption{GPU hazard ratios from Cox regression model on {\tt new}
    batch.}
  \label{fig:cox-new}
\end{figure}

We find that the average hazard ratios strongly correlate with
detailed nuances of the system cooling architecture.  The figures give
hazard ratios for each of the location variables, giving the ratio for
each level compared to its first (0) level (consequently the 0 level
is always 1). Due to the exponential nature of the CPH model, the
horizontal hazard ratio axis is on a log scale. The estimates include
uncertainty, which too is most reliably computed on a log scale (see,
for example, \cite{Ostrouchov88}). In the figures, we also include the
number of units, N, at risk and the number of events that occurred in
each category. All the factors ({\em col, row, cage, slot, node}) are
balanced with respect to the number of units at risk and consequently
nearly orthogonal (each level of a factor contains all levels of the
other factors) providing an almost ``designed'' experiment nature to
this analysis.

Our interpretation of correlation with detailed nuances of the cooling
system comes mostly from the hazard ratios for the old batch. The new
batch has mot had many failure events and the uncertainty bars of
nearly all ratios include 1, which is no difference. In the old batch,
we see that {\em cage} has the strongest effect, putting the highest
hazard ratio on cage 2, which is consistent with lowest survivals in
the KM analysis earlier. Its ratio value near 20 has to be interpreted
with caution because it is a time averaged value on a log scale. This
aspect is not captured by the uncertainty, which accounts for the
randomness of the failures and not for the geometric time averaging
\cite{coxhazardinterpret}.

>>>>

If $T$ is the random variable of a GPU fail time, then its cumulative
distribution function $F(t) = Pr\{T < t\}$ gives the probability that
a GPU fails by duration $t$. The survival function is its complement
\begin{displaymath}
  S(t) = Pr\{T \geq t\} = 1 - F(t).
\end{displaymath}
It is the probability of being operational at duration $t$.  We use
the R packages \pkg{survival} and \pkg{survminer} for the KM analysis,
which is reported in Figure~\ref{fig:km-all-cage-node}. Within each
{\em batch}, separate survival curves are computed for each {\em cage}
by {\em node} combination. Along with the survival curve estimate,
this analysis provides 95\% confidence region for survival probability
shaded around the curves.
% Inserting figure here so it appears at top of next page
\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{figs/cox_o001.pdf}
  \caption{GPU hazard ratios from Cox regression model on {\tt old}
    batch.}
  \label{fig:cox-old}
\end{figure}
\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{figs/cox_n001.pdf}
  \caption{GPU hazard ratios from Cox regression model on {\tt new}
    batch.}
  \label{fig:cox-new}
\end{figure}

@@

CPH is considered a semi-parametric model as there are no assumptions
about the shape of the baseline hazard function. Its strongest
assumption is that the hazards are proportional. There are a number of
tests for this, including graphical diagnostics in the \pkg{survminer}
package as well as checking that survival curves for categorical
covariates do not cross. We ran these diagnostics and concluded that
the hazards for our location categories are approximately
proportional. However, survival functions partitioned on {\em batch}
do cross and so we fit the model to the new and the old batches
separately. The results are presented in Figures~\ref{fig:cox-old}
and~\ref{fig:cox-new}.

We find that the average hazard ratios strongly correlate with
detailed nuances of the system cooling architecture.  The figures give
hazard ratios for each of the location variables, giving the ratio for
each level compared to its first (0) level (consequently the 0 level
is always 1). Due to the exponential nature of the CPH model, the
horizontal hazard ratio axis is on a log scale. The estimates include
uncertainty, which too is most reliably computed on a log scale (see,
for example, \cite{Ostrouchov88}). In the figures, we also include the
number of units, N, at risk and the number of events that occurred in
each category.

All the factors ({\em col, row, cage, slot, node}) in the old batch
are balanced with respect to the number of units at risk and
consequently nearly orthogonal (each level of a factor contains all
levels of the other factors) seemingly an almost ``designed''
experiment nature to this analysis. This is not the case for the new
batch, where the cage levels have very different numbers of units at
risk. But this also points out that even for the old batch the balance
holds only at the outset and as life proceeds, failing units are
replaced with new units and the balance degrades.

Our interpretation of correlation with details of the cooling system
comes mostly from the hazard ratios for the old batch. The new batch
has mot had many failure events and the uncertainty bars of nearly all
ratios include 1, which is no difference. In the old batch, we see
that {\em cage} has the strongest effect, putting the highest hazard
ratio on cage 2, which is consistent with lowest survivals in the KM
analysis earlier. Its ratio value near 20 has to be interpreted with
caution because it is a time averaged value on a log scale and suffers
from the degrading balance mentioned in the previous paragraph. This
aspect is not captured by the uncertainty, which accounts for the
randomness of the failures and not for the geometric time averaging
\cite{coxhazardinterpret}. Consequently, we interpret the pattern of
relative hazards rather than actual ratio magnitudes. More in-depth
analysis with penalized estimation methods like \cite{bender2019} can
take such balance issues in the exposure history into account.
